Pathways to Early Vocabulary Acquisition

### 2024-12-03 other/datasheet/datasheet_wordbank.qmd
Typo updated.

---
title: "Datasheet for Wordbank"
subtitle: "A Datasheet Used to Understand the Study on  'Pathways to Early Vocabulary Acquisition'"
author: Yongqi Liu
thanks: "Code and data are available at: [Early Vocabulary Acquisition Pathway](https://github.com/Cassieliu77/Early_Vocabulary_Acquisition_Pathway.git)"
date: today
abstract: "The Wordbank dataset is a comprehensive, publicly accessible resource designed to advance the study of early vocabulary acquisition in children aged 16–30 months. It aggregates anonymized data from the MacArthur-Bates Communicative Development Inventories (CDIs), spanning 29 languages and dialects, to support research on linguistic development, cross-linguistic comparisons, and developmental trends. Maintained by Stanford University’s Language and Cognition Lab, the dataset features both norming and non-norming samples, offering robust analytical capabilities while capturing the diversity of linguistic environments. This datasheet provides an in-depth overview of the dataset’s purpose, structure, collection methodology, distribution, and ethical considerations, highlighting its utility as a transparent and versatile tool for research and education."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The Wordbank dataset was created to provide a centralized platform for archiving, sharing, and analyzing anonymized data from the MacArthur-Bates Communicative Development Inventories (CDI). Its primary objective is to support researchers in studying the cognitive development processes of children during early childhood. By compiling data across multiple languages, the dataset serves as a comprehensive resource for both norming studies and diverse contributions from various research projects. It addresses the need for an accessible repository of CDI data, enabling researchers to generate norms for populations and individual vocabulary items interactively. It also bridges a critical gap by facilitating cross-linguistic comparisons of early lexical development and offering standardized benchmarks alongside diverse linguistic samples. Building on its predecessor, CLEX, Wordbank enhances functionality as both a research and teaching tool. It allows users to analyze developmental patterns, customize norms, and conduct interdisciplinary studies, making it an invaluable asset in linguistic, psychological, and developmental research.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The Wordbank dataset was developed by the Language and Cognition Lab at Stanford University, under the direction of Dr. Michael C. Frank. This initiative is part of Stanford’s broader commitment to advancing research in child language acquisition and cognitive development. The platform compiles data from the MacArthur-Bates Communicative Development Inventories (CDIs) across multiple languages, serving as a valuable resource for researchers worldwide.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The development of Wordbank was funded by a grant from the National Science Foundatio as well as generous support from the MB-CDI Advisory Board.
4. *Any other comments?*
    - The dataset is a crucial resource for advancing research on children’s cognitive and linguistic development. It highlights critical windows of vocabulary growth and integrates both norming and non-norming samples, providing standardized benchmarks alongside diverse linguistic contexts. With its cross-linguistic breadth and detailed lexical categorization, it serves as a powerful tool for analyzing global patterns in early vocabulary acquisition and fostering a deeper understanding of developmental trajectories.

@@

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- The instances in the dataset are children aged 16–30 months and their parents' responses from the MacArthur-Bates Communicative Development Inventories (CDIs). Each instance is tied to a specific child and consists of detailed information about the words they understand (comprehension) and can use (production), categorized into lexical groups such as nouns, verbs, and adjectives.
2. *How many instances are there in total (of each type, if appropriate)?*
	- Wordbank now contains data from 92,771 children and 105,290 CDI administrations, across 42 languages and 89 instruments.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- No. The wordbank package collected a sample drawn from responses collected using the MacArthur-Bates Communicative Development Inventories (CDIs) across various languages and studies. It includes both norming samples, which are standardized and representative of general language acquisition trends, and non-norming samples, contributed by researchers studying specific subpopulations or contexts. While the norming data provide a validated baseline for comparison, the non-norming data add diversity but may not fully represent population-level trends.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	Here is a single type of instance in the dataset, but each instance includes multiple dimensions, such as:

@@

	- The dataset is self-contained but derived from the CDI and Wordbank repositories. External tools like Wordbank R package were used for data extraction and processing.

@@

	- Yes. Individuals (typically parents or guardians of the children) were notified about the data collection. Notification was not seen in the provided survey form from the MacArthur-Bates Communicative Development Inventories, but it may exist in other real-person communications and records. The consent process included clear communication about the purpose of the study, how the data would be used, and assurances of anonymity. This information was typically conveyed through a consent form or verbal explanation, depending on the study protocol. 

@@

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- Yes. The dataset has been utilized to investigate vocabulary acquisition patterns in children aged 16–30 months. Specifically, it has been used to: 
	i). Predicting high vocabulary proficiency based on variables like age, norming status, and lexical categories.
	ii). Examining patterns in the development of different word categories, such as Function Words and Sensory Words, to understand their relative acquisition timelines.
	iii). Investigating the impact of norming and non-norming environments on language development to identify contextual factors influencing vocabulary growth.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- The dataset is linked to studies and resources available on the CDI Wordbank platform, as well as to related papers focused on early language development. Future research using this dataset will reference these repositories as part of their methodology. The Wordbank repository, which includes tools for accessing and analyzing the dataset, is accessible on GitHub at: https://github.com/langcog/wordbankr.
3. *What (other) tasks could the dataset be used for?*
	- Exploration of Language Developmental Milestones: Studying the progression of specific word categories over time.

@@

	- Yes. The dataset will be updated periodically. The number of CDIs stored in Wordbank is continually growing, and the development team are always interested in adding more datasets to the site. To see which researchers have already contributed, click here: [Wordbank contributors](http://wordbank.stanford.edu/contributors). If you are interested in contributing your own data, please contact via: wordbank-contact@stanford.edu.

>>>>

---
title: "Datasheet for Wordbank"
subtitle: "A Datasheet Used to Understand the Study on  'Pathways to Early Vocabulary Acquisition'"
author: Yongqi Liu
thanks: "Code and data are available at: [Early Vocabulary Acquisition Pathway](https://github.com/Cassieliu77/Early_Vocabulary_Acquisition_Pathway.git)"
date: today
abstract: "The Wordbank dataset is a comprehensive, publicly accessible resource designed to advance the study of early vocabulary acquisition in children aged 16–30 months. It aggregates anonymized data from the MacArthur-Bates Communicative Development Inventories (MB-CDIs), spanning 29 languages and dialects, to support research on linguistic development, cross-linguistic comparisons, and developmental trends. Maintained by Stanford University’s Language and Cognition Lab, the dataset features both norming and non-norming samples, offering robust analytical capabilities while capturing the diversity of linguistic environments. This datasheet provides an in-depth overview of the dataset’s purpose, structure, collection methodology, distribution, and ethical considerations, highlighting its utility as a transparent and versatile tool for research and education."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The Wordbank dataset was created to provide a centralized platform for archiving, sharing, and analyzing anonymized data from the @cdi. Its primary objective is to support researchers in studying the cognitive development processes of children during early childhood. By compiling data across multiple languages, the dataset serves as a comprehensive resource for both norming studies and diverse contributions from various research projects. It addresses the need for an accessible repository of CDI data, enabling researchers to generate norms for populations and individual vocabulary items interactively. It also bridges a critical gap by facilitating cross-linguistic comparisons of early lexical development and offering standardized benchmarks alongside diverse linguistic samples. Building on its predecessor, CLEX, Wordbank enhances functionality as both a research and teaching tool. It allows users to analyze developmental patterns, customize norms, and conduct interdisciplinary studies, making it an invaluable asset in linguistic, psychological, and developmental research.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The Wordbank dataset was developed by the Language and Cognition Lab at Stanford University, under the direction of Dr. Michael C. Frank. This initiative is part of Stanford’s broader commitment to advancing research in child language acquisition and cognitive development. The platform archives data from the @cdi, across multiple languages, serving as a valuable resource for researchers worldwide.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The development of Wordbank was funded by a grant from the National Science Foundatio as well as generous support from the MB-CDIs Advisory Board.
4. *Any other comments?*
    - The dataset is a crucial resource for advancing research on children’s cognitive and linguistic development. It highlights critical windows of vocabulary growth and integrates both norming and non-norming samples, providing standardized benchmarks alongside diverse linguistic contexts. With its cross-linguistic breadth and detailed lexical categorization, it serves as a powerful tool for analyzing global patterns in early vocabulary acquisition and fostering a deeper understanding of developmental trajectories.

@@

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- The instances in the dataset are children aged 16–30 months and their parents' responses from the @cdi, Each instance is tied to a specific child and consists of detailed information about the words they understand (comprehension) and can use (production), categorized into lexical groups such as nouns, verbs, and adjectives.
2. *How many instances are there in total (of each type, if appropriate)?*
	- Wordbank now contains data from 92,771 children and 105,290 CDI administrations, across 42 languages and 89 instruments.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- No. The wordbank package collected a sample drawn from responses collected using the @cdi, across various languages and studies. It includes both norming samples, which are standardized and representative of general language acquisition trends, and non-norming samples, contributed by researchers studying specific subpopulations or contexts. While the norming data provide a validated baseline for comparison, the non-norming data add diversity but may not fully represent population-level trends.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	Here is a single type of instance in the dataset, but each instance includes multiple dimensions, such as:

@@

	- The dataset is self-contained but derived from the MB-CDIs website and Wordbank repositories. External tools like Wordbank R package were used for data extraction and processing.

@@

	- Yes. Individuals (typically parents or guardians of the children) were notified about the data collection. Notification was not seen in the public-accessed survey from the MB-CDI website, but it may exist in other real-person communications and records. The consent process included clear communication about the purpose of the study, how the data would be used, and assurances of anonymity. This information was typically conveyed through a consent form or verbal explanation, depending on the study protocol. 

@@

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- Yes. The dataset has been utilized to investigate vocabulary acquisition patterns in children aged 16–30 months. Specifically, it has been used to: 
	i). Predicting high vocabulary proficiency based on variables like age, norming status, and lexical categories.
	ii). Examining patterns in the development of different word categories, such as Function Words and Sensory Words, to understand their relative acquisition timelines.
	iii). Investigating the impact of norming and non-norming environments on language development to identify contextual factors influencing vocabulary growth.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- The dataset is linked to studies and resources available on the Wordbank platform, as well as to related papers focused on early language development. Future research using this dataset will reference these repositories as part of their methodology. The Wordbank repository, which includes tools for accessing and analyzing the dataset, is accessible on GitHub at: https://github.com/langcog/wordbankr.
3. *What (other) tasks could the dataset be used for?*
	- Exploration of Language Developmental Milestones: Studying the progression of specific word categories over time.

@@

	- Yes. The dataset will be updated periodically. The number of CDIs stored in Wordbank is continually growing, and the development team are always interested in adding more datasets to the site. To see which researchers have already contributed, click here: [Wordbank Contributors](http://wordbank.stanford.edu/contributors). If you are interested in contributing your own data, please contact via: wordbank-contact@stanford.edu.

### 2024-12-03 paper/paper.qmd
Typo updated.

---
title: "Pathways to Early Vocabulary Acquisition"
subtitle: "Mapping English Language Development Trajectories in Children Aged 16–30 Months"
author: Yongqi Liu
thanks: "Code and data are available at: [Early Vocabulary Acquisition Pathway](https://github.com/Cassieliu77/Early_Vocabulary_Acquisition_Pathway.git)"
date: today
date-format: long
abstract: "This study examines vocabulary acquisition in children aged 16–30 months using data from the Wordbank and MacArthur-Bates Communicative Development Inventories. Age is the strongest predictor, with vocabulary size increasing consistently and foundational categories like Function Words, Living Things, and Objects showing higher predicted probabilities, indicating their universal acquisition. Complex categories such as Sensory Words, Adjectives, and Places exhibit broader distributions and lower probabilities, reflecting their reliance on contextual understanding and cognitive complexity. These findings underscore age-specific patterns in vocabulary growth and the need for targeted interventions to support diverse word categories learning."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

The original dataset was obtained from @cdi and downloaded using the R package @wordbankr. After a thorough cleaning process such as grouping related items and removing missing values, the analysis focuses on key variables, including category, age, comprehension, production and is_norming. These variables form the foundation of the analysis dataset, with an overview provided in @tbl-data.

@@

Distinct peaks in the dataset, such as at 24 and 25 months, may indicate intentional focus points for testing or align with developmental benchmarks tied to standardized assessments like the MacArthur-Bates Communicative Development Inventories (CDI). However, this uneven distribution also highlights the challenges of assessing vocabulary development in younger children, where verbal communication is less developed, and parental reporting is more variable. These disparities underscore the importance of standardizing age in statistical models to ensure accurate analysis and interpretation, while also exercising caution when generalizing findings to underrepresented age groups.

@@

This study sheds light on key aspects of vocabulary acquisition and offers meaningful directions, but certain limitations must be acknowledged. Firstly, the reliance on parental reports, as collected through the MacArthur-Bates Communicative Development Inventories (CDI), introduces potential biases. Parents may overestimate or underestimate their child’s abilities due to memory limitations, subjective interpretation, or social desirability. Such biases could affect the accuracy of the data, particularly for categories requiring nuanced understanding, like Sensory Words. To mitigate these limitations, future research should incorporate complementary methods, such as direct observational studies or experimental assessments, to provide more reliable and objective measurements of children’s vocabulary development.

@@

The MacArthur-Bates Communicative Development Inventories (CDI) form the backbone of this study, offering a well-validated framework for capturing early vocabulary acquisition. The CDI consists of structured checklists where parents can report their child’s comprehension and production of specific words. These items are categorized into groups, such as nouns, verbs, and predicates, enabling researchers to analyze patterns across different lexical categories. Despite its strengths, the reliance on parental reports introduces variability in accuracy, particularly for some abstract word categories such as sensory, where interpretation can differ among respondents.

>>>>

---
title: "Pathways to Early Vocabulary Acquisition"
subtitle: "Mapping English Language Development Trajectories in Children Aged 16–30 Months"
author: Yongqi Liu
thanks: "Code and data are available at: [Early Vocabulary Acquisition Pathway](https://github.com/Cassieliu77/Early_Vocabulary_Acquisition_Pathway.git)"
date: today
date-format: long
abstract: "This study examines vocabulary acquisition in children aged 16–30 months using data from the Wordbank and MacArthur-Bates Communicative Development Inventories (MB-CDIs). Age is the strongest predictor, with vocabulary size increasing consistently and foundational categories like Function Words, Living Things, and Objects showing higher predicted probabilities, indicating their universal acquisition. Complex categories such as Sensory Words, Adjectives, and Places exhibit broader distributions and lower probabilities, reflecting their reliance on contextual understanding and cognitive complexity. These findings underscore age-specific patterns in vocabulary growth and the need for targeted interventions to support diverse word categories learning."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

The original dataset was obtained from @cdi and downloaded using the R package from @wordbankr. After a thorough cleaning process such as grouping related items and removing missing values, the analysis focuses on key variables, including category, age, comprehension, production and is_norming. These variables form the foundation of the analysis dataset, with an overview provided in @tbl-data.

@@

Distinct peaks in the dataset, such as at 24 and 25 months, may indicate intentional focus points for testing or align with developmental benchmarks tied to standardized assessments. However, this uneven distribution also highlights the challenges of assessing vocabulary development in younger children, where verbal communication is less developed, and parental reporting is more variable. These disparities underscore the importance of standardizing age in statistical models to ensure accurate analysis and interpretation, while also exercising caution when generalizing findings to underrepresented age groups.

@@

This study sheds light on key aspects of vocabulary acquisition and offers meaningful directions, but certain limitations must be acknowledged. Firstly, the reliance on parental reports, as collected through the MacArthur-Bates Communicative Development Inventories (MB-CDIs), introduces potential biases. Parents may overestimate or underestimate their child’s abilities due to memory limitations, subjective interpretation, or social desirability. Such biases could affect the accuracy of the data, particularly for categories requiring nuanced understanding, like Sensory Words. To mitigate these limitations, future research should incorporate complementary methods, such as direct observational studies or experimental assessments, to provide more reliable and objective measurements of children’s vocabulary development.

@@

The MacArthur-Bates Communicative Development Inventories (MB-CDIs) form the backbone of this study, offering a well-validated framework for capturing early vocabulary acquisition. The CDI consists of structured checklists where parents can report their child’s comprehension and production of specific words. These items are categorized into groups, such as nouns, verbs, and predicates, enabling researchers to analyze patterns across different lexical categories. Despite its strengths, the reliance on parental reports introduces variability in accuracy, particularly for some abstract word categories such as sensory, where interpretation can differ among respondents.
